---
title: "Joining Assignment"
author: "Your Name Here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

```

# Load the datasets

```{r}

counties_demographics <- readRDS(here("data", "counties_demographics.rds"))
counties_elexresults <- readRDS(here("data", "counties_elexresults.rds"))
counties_rural_urban_suburban <- readRDS(here("data", "counties_rural_urban_suburban.rds"))



```

# Exercises

1)  

Join together the election results and urban/suburban/rural tables above, and then use the merged table to:

-   filter for just Virginia

-   group_by and summarize to calculate the mean pct of Trump 2020 vote (`pct_r_2020`) by urban/rural categories (`rural_urban`)

```{r}

merged_table <- left_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid")

virginia_table <- filter(merged_table, state_code == "VA")

virginia_table %>%
  group_by(rural_urban) %>%
  summarize(mean_pct_r_2020 = mean(pct_r_2020))


```

2)  

Using the joined election results / urban/suburban/rural table created above:

-   filter for just Virginia

-   count up how many counties were won by party in 2020 (`winner_2020`) by urban/rural categories

```{r}

virginia_table %>%
  group_by(rural_urban, winner_2020) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = winner_2020, values_from = count) %>%
  replace(is.na(.), 0)


```

3)  

Join together the election results and demographics tables above, and then use the joined table to:

-   filter for just Florida

-   group/summarize to calculate the mean pct of Trump 2020 vote in the counties above or below national average for college graduates (`pct_ed_college_all_abovebelow_natl`)

```{r}

merged_table1 <- left_join(counties_elexresults, counties_demographics, by = "pct_d_2020")

florida_table <- filter(merged_table1, state_code == "FL")


national_avg_college <- mean(florida_table$pct_ed_college_all)

florida_table %>%
  mutate(college_abovebelow_natl = ifelse(pct_ed_college_all > national_avg_college, "Above national average", "Below national average")) %>%
  group_by(college_abovebelow_natl) %>%
  summarize(mean_pct_r_2020 = mean(pct_r_2020.x))


```

4)  

Using the joined table you made above:

-   group/summarize to calculate the mean pct of Trump 2020 vote by counties above or below the national average for college graduates (`pct_ed_college_all_abovebelow_natl`).

    Note we're looking for the entire country here.

```{r}

national_avg_college <- mean(merged_table1$pct_ed_college_all)

merged_table1 %>%
  mutate(college_abovebelow_natl = ifelse(pct_ed_college_all > national_avg_college, "Above national average", "Below national average")) %>%
  group_by(college_abovebelow_natl) %>%
  summarize(mean_pct_r_2020 = mean(pct_r_2020.x))


```

Now using the joined datasets above, come up with two groupings you find interesting or newsworthy in a particular state, and replicate the steps above them them instead.

5)  

```{r}
#This table groups counties in Pennsylvania by their rural-urban status and calculates the mean percentage of the Biden 2020 vote and the Clinton 2016 vote in each group. 

merged_table3 <- left_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid")

pennsylvania_table <- filter(merged_table3, state_code == "PA")

pennsylvania_table %>%
  group_by(rural_urban) %>%
  summarize(mean_pct_d_2020 = mean(pct_d_2020),
            mean_pct_d_2016 = mean(pct_d_2016))




```

6)  

```{r}

# This groups counties in Pennsylvania by their rural-urban status and calculates the mean change in percentage vote share for Trump and Biden between 2016 and 2020 in each group. 

pennsylvania_table %>%
  group_by(rural_urban) %>%
  summarize(mean_pct_r_change = mean(pct_r_2020 - pct_r_2016),
            mean_pct_d_change = mean(pct_d_2020 - pct_d_2016))





```

7)  Extra credit

Up until now we've joined two tables together. But we actually have three tables in this project. How might we join all three together into a single table, so all the measures for each county are combined in one place? (Hint: you can do this is multiple steps, doesn't have to be one step only.)

```{r}


merged_table2 <- left_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid")

merged_table4 <- left_join(merged_table1, counties_demographics, by = "geoid")



```
